Solar Flares Across the Spectrum: Observations and Instrumentation

### 2025-04-11 introduction/intro_2.tex
Local changes

The Sun is our nearest astrophysical object, which serves as a natural laboratory for various branches of Physics - atomic physics, spectroscopy, turbulence, magneto-hydrodynamics, stellar evolution, dynamo and magnetism and plays a key role in space and terrestrial weather. The Sun is also the primary source of light and energy on Earth. So it is only natural that we try to understand the different physical processes involved in the energetics of the Sun, which directly affects life on Earth. In this chapter, I briefly introduce the Sun, various layers of its atmosphere and Solar flares. Finally, we provide a motivation for the thesis and a brief outline.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The Sun}\label{Sun}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The Sun is a G2V star, with surface luminosity $3.86~\times~10^{26}$~W and an effective temperature of $\sim 5780~K$. The Sun is mostly made out of Hydrogen (92.1\%) and Helium (7.8\%), along with negligible quantities of heavier elements C, N, O, Mg, Si, Ne and Fe, among others. The study of the Sun can be divided into three components, \textit{viz.} the Solar interior, the Solar atmosphere and the Heliosphere. Fig.~\ref{fig_solar_int} shows a schematic diagram of various regions of the Sun. Below, we briefly describe these three regions.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The Solar interior}\label{solar_int} 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\info{First you have to say what is the interior! For example the interior is made of three regions -- the core, the radiative zone and the convective zone. The innermost zone, i.e., the core}
Depending on various physical characteristics, the Solar interior can be divided into three zones {\it e.g.} the core, radiative zone and convective zone. In the innermost zone, the core, the Sun has a temperature of about 15~MK and a density of $\mathrm{1.6\times 10^{5}~km.m^{-3}}$. Here Hydrogen fuses into Helium primarily by p-p chain reaction \citep{bethe38} and to some extent by CNO cycle \citep{bethe39}. Beyond the core, we have the optically thick radiative zone, where the Hydrogen and Helium are completely ionized due to very high density. The high energy $\gamma$-ray photons generated in the core have a very small mean free path in this region, as they collide numerous times and propagate across this region. As we move radially outwards, temperature and density both decrease slowly due to gravitational equilibrium. Due to this reduction of temperature, there is a decrease in the thermal motion of the particles, and the hydrogen and helium start to recombine, giving rise to an increase in opacity at the top of the radiative zone. This change in opacity makes the propagation of photons via radiation less efficient, as photons get trapped and reabsorbed more. This creates a sharp temperature gradient due to the "photon pileup". This temperature gradient gives rise "convective instability". This is the convective zone where the energy from the inner layers of the Sun is transported adiabatically to the surface of the Sun. 
%In between the radiative and the convective zone, there is \sroy{believed to be }a thin layer, roughly at $\sim~0.7~R_{\odot}$ known as the tachocline. \sroy{This region} is believed to play a major role in generating Sun's magnetic field via the Dynamo process \citep{corbard01}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The Solar atmosphere}\label{solar_atmos}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The layers outside the convection zone together form the Solar atmosphere and is further divided into different regions. The photosphere, which is also known as the Sun's surface, the chromosphere, the transition region and the corona. While we can define geometrical height or depths to define these layers, it is much more useful to define them using the local optical depths of various spectral lines, which are directly used to observe the layers. This also gives us an idea about which spectral features originate mainly from which portions of the Solar atmosphere. Fig.~\ref{fig_solar_atm} shows the variation of temperature and number density across various layers of the Solar atmosphere \citep{philips08}.

%% ############# %%
\begin{figure}[ht!]
    \centering
    \includegraphics[width = 0.8\linewidth]{Figures/solar_atm.png}
    \caption[Density and temperature profiles across various layers of the Sun.]{The variation of electron temperature (solid line) and electron density ($\mathrm{N_{e}}$, dashed line) and density of neutral hydrogen atoms ($\mathrm{N_{H}}$, dot-dashed line) in the Solar atmosphere \citep{philips08}.}
    \label{fig_solar_atm}
\end{figure}
%% ############# %%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{The photosphere}\label{photosphere}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The photosphere is defined to be the layer where $\tau_{\lambda~\sim~5500}=1$. This is in the green part of the visible spectrum and the Sun is opaque in visible beyond this layer, hence it is called to be the surface of the Sun. This layer is 400{--}600~km thick and has an effective temperature of 5780 K. The magnetic field lines arising from the tachocline penetrate the photosphere and create a `carpet' over the whole region \citep{priest14}. As mentioned earlier, this is the Solar surface known as the `quiet Sun'(QS). This region exhibits an average magnetic flux density of 10 {--} 50 G. The QS surface is covered with cells of roughly four sizes, namely, granules, mesogranules, supergranules and giant cells.

Some regions exhibit much stronger magnetic flux density, often associated with highly twisted magnetic field structures. These magnetic features are manifested as Sunspots. The Sunspots harbour some of the strongest concentrated magnetic fields, reaching up to several kilogauss in strength. The other noteworthy Solar magnetic structure apart from these are believed to be spatially unresolved flux tubes composing a network with field strengths $\sim$ {\it $O(10^{3} G)$}\citep{solanki93,grossman96,rubio19}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{The chromosphere}\label{chromosphere}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The chromosphere is a highly non-uniform dynamic layer with a thickness of 1500 {--} 3000 km, with increasing temperature (up to $10^{4}$ K) and decreasing number density. As seen in Fig.~\ref{fig_solar_atm}, the temperature of the chromosphere saturates before the dramatic rise in the transition region. This saturation is attributed to a steady deposition of acoustic energy by the creation of shock waves \citep{carlsson07}. The chromosphere also exhibits a sharp gradient in the plasma $\beta$ factor, non-LTE conditions, and the dominance of wave motions. The observable features in the chromosphere are active regions, sunspots, plages and also light bridges as labelled in images taken by IRIS SJI (slit-jaw imager) 2796 {\AA} (left panel) and 2832 {\AA} (right panel) displayed in Fig.~\ref{fig:sji_features}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{The corona}\label{corona}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The outermost layer of the atmosphere is the corona. The corona maintains a temperature of 1-2~MK during ambient conditions but can rise up to tens of MK during explosive eruptions such as Solar flares. Due to such high temperatures, the coronal gas is fully ionized. It is only visible to the naked eye during the total Solar eclipse or with coronagraphs due to Thomson scattering of the photosphere with free electrons in the corona. corona also emits in various emission lines, e.g. Mg, Fe, Si, S, Ca, O, etc., across a wide band of electromagnetic spectrum ranging from X-ray to visible wavelengths. Similar to the other layers, the corona exhibits a variety of structures such as ARs, coronal holes, bright points, filaments, and flares. The flare arcades are most prominently visible in the corona. %\info{I made a lot of changes in this section. please see if all is consistent.}

@@

Solar flares are the most powerful energetic explosions in the Solar system. They are described as a sudden increase in brightness in localized areas of the Sun. Within tens of minutes, they can release over $10^{32}$ erg of energy, which is emitted across the entire electromagnetic spectrum, i.e., from radio to gamma rays. They can also launch highly energetic particles into the interplanetary medium. Flares occur in magnetic active regions. It has been estimated that the amount of flare energy released is comparable to the free energy stored in the magnetic system \citep{emslie12,ash17}. %The term "flare" is generally used explicitly for the entire magnetically-driven event's electromagnetic radiation, as it is the most significant fraction of the total energy liberated. 
The total energy released varies from event to event. It is also known that larger events occur much less frequently than smaller events.

Fig.~\ref{fig:sji_features} displays one of the most well-known and well-studied Solar eruptions from Aug 31st, 2012. During this event, an associated giant prominence erupted from the southeast limb, which was visible in all AIA channels as shown in Fig.~\ref{fig:aia-feature}. The sunspots were also clearly visible and labelled in the HMI continuum (see Fig.~\ref{fig:aia-feature}h). In Fig.~\ref{fig:aia-feature}i, we show the co-aligned AIA 304 {\AA} and HMI continuum observation. The Sunspots lie at the base of the erupted prominence, accompanied by the hot thermal plasma associated with the accompanying flare (visible in Fig.~\ref{fig:aia-feature}e {--} f). The novelty of such observation lies in the clear spatial connection across various layers of the Sun, connected via energy and momentum transported from one layer to the other.

@@

On September 1, 1859, R.C. Carrington and R. Hodgson independently observed the first flare in white-light continuum~\citep{carrington1859,hodgson1859}. Such localized brightening on the Sun has remained an enigma ever since. 
Shortly after the observation by Carrington and Hodgson, the Sun was being studied extensively in the H$\alpha$ wavelength, which essentially images the chromosphere, and the reports of flaring events became increasingly frequent and progressively more complex. No two events were similar, as there were variations observed in the size of the source, ejections of plasma, and shock waves driving into the interplanetary space. Advances in radio technology during the Second World War affirmed the detection of the presence of non-thermal electrons in the Solar corona during military radar operations \citep{hey46}. Around the same time, S.E. Forbrush noticed ground-level cosmic ray enhancement during major Solar flares. These discoveries eluded that the flaring events do not only involve the thermal plasma but somehow connect with high energy particles and involve the corona.

In the late 1950s, using rockets and balloons, the Sun was observed with hard X-rays ($\ge~10~keV$). \cite{peterson59} discovered the first hard X-ray emission during a flare in 1958. Later on, it was deduced from the enhancements observed in radio and hard X-rays that the ejected energetic particles may contain a substantial fraction of the initial energy released \citep{brown71}. Note that while the hard X-ray is created by the bremsstrahlung radiation of the electrons colliding into dense material, resulting in a power-law energy distribution, the broadband radio emission from 1 to 100 GHz is created from gyrosynchrotron emission. Flares are now routinely observed with different space-borne instruments, in particular in the EUV and soft X-ray($\le$~10~keV), which shows that the energy released from the flare heats the local plasma to temperatures beyond 30~MK.

@@

\noindent and is known as the ``Neupert effect''. This relationship shows that the soft X-ray radiation primarily originates from a thermal plasma which is heated by the energy of the flaring event deposited by the accelerated electrons. We note, however, that Eqn.\ref{npt_eff_eq} is only valid if cooling by conduction is negligible and/or the role of ions in the flare energy deposition is minimal \citep{veronig05}.

%% ################################################################# %%
\subsection{``Standard'' model of Solar flare}\label{sol_flr_std_mod}
%% ################################################################# %%

Based on various multi-wavelength studies of Solar flares, combined with theoretical modelling a `standard' model of Solar flares has been advocated by \cite{carmichael64,sturrock66,hirayama74,kopp76}, and termed as \textsl{CSHKP} model. The flares occur due to the abrupt release of magnetic free energy, which was previously stored in the coronal magnetic field via flux emergence and surface motions \citep{forbes06}. The majority of the strongest flares are eruptive in nature. The `standard' model attributes the release of flare energy to magnetic reconnection in the corona, which happens due to the eruption and/or coronal mass ejection \citep{shibata95,lin_n_forbes00,moore01,priest_n_forbes02}. It is noteworthy that, despite continuous attempts over the decades, this is not at all a quantitative model. Rather, it is an attempt to unify the multi-wavelength observations and modelling attempts in an orderly manner. Naturally, the model can not explain all the observations from various flares. A schematic picture of the model is shown in Fig.~\ref{fig:std_mod}. Below we describe the salient points of the `standard' model:

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[ht!]
    \centering
    \hspace{5cm}
    \includegraphics[trim={1cm 8cm 0cm 8cm}, clip, width=0.95\linewidth]{Figures/std_mod.pdf}
    \caption{A schematic diagram of the standard model of Solar flare.}
    \label{fig:std_mod}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%
\begin{enumerate}
    \item In the standard model, following the reconnection event occurring that releases the magnetic free energy, electrons and ions from the reconnection site are accelerated along the realigned magnetic field lines downwards. In the case of an associated CME, parts of the magnetic energy is deposited as kinetic energy. The coronal reconnection source where the energy is released is often observed in HXR. The loop-top HXR sources are often accompanied by the thermal SXR sources. The loop top HXR source is observed above the coronal SXR source \citep{masuda94}.

    \item The precipitated particles move along the field lines downwards along the magnetic field lines, giving rise to the microwave and radio observations seen from flares via gyro magnetic radiation at various frequencies dictated by the characteristics of local plasma.

    \item The accelerated particles hit the chromosphere, which is considerably denser than the corona, giving rise to Hard X-rays and gamma-rays observed from the foot points. This also explains why the coronal hard X-ray source is considerably softer than the foot points. The energetic particles deposit their energy into the local chromosphere as they go through a series of collisions and eventually thermalize \citep{brown83}. Foot point sources are occasionally detected on H-alpha flare ribbons that are approximately 30,000 km apart, marking the base of a magnetic arcade \citep{masuda01}.

    \item $\gamma$-ray emission is visible between 0.8 {--} 20 MeV generated by the accelerated ions. The neutron capture line at 2.2 MeV forming Deuterium is usually the strongest characteristic line visible from such emissions. Protons and other accelerated ions collide with the dense chromosphere and produce neutrons. These neutrons are captured by ambient protons to form Deuteron \citep{ramtay74,hua87}. There were later RHESSI observations \citep{hurford_2003,hurford_2006} that showed that the $\gamma$-ray foot-points did not coincide with the HXR foot-points. This demonstrated that the ions and electrons were accelerated differently, as proposed by \cite{emslie_2004}.

    \item  The thermalized plasma emits in thermal bremsstrahlung, giving rise to soft X-ray and EUV radiation. The heating of the local chromosphere gradually increases the local pressure. When the growing pressure gradient builds up enough, the local plasma starts expanding upwards and slowly fills up the coronal loops with soft X-ray-emitting plasma. This phenomenon is known as "Chromospheric evaporation". This was first reported by \cite{doscheck80,feldman80}. \cite{antonucci89} reported velocities of 300 {--} 400 km/s with temperature $\sim$ 20 MK in \ion{Ca}{19}, while later on \cite{milligan06} reported much gentler up-flow with velocities $\sim$ 200 km/s in \ion{Fe}{19} line from {\it SOHO}/CDS observations. Several simulations also reproduced evaporation arising from non-thermal particle precipitation \citep{sterling93,hori98,reeves07}.
\end{enumerate}
%%%%%%%%%%%%%%%%%

It is important to emphasise that while the standard model attempts to explain and unify various kinds of differences seen from the numerous flares we observe, there are several cases where the standard model cannot explain the observations. For example, there have been observations of flares where the hard X-ray footpoint does not form. It is proposed that in these cases, the plasma in the coronal loops is so dense that the accelerated particles pass through enough collisions to thermalize within the flare loops before reaching the chromosphere. In this process, they may defuse the energy more evenly within the loop rather than impulsively dumping it at the base of the loops \citep{veronig02,veronig04}. Another flare which previously occurred at the same region might explain the denser coronal loops \citep{strong84,bone07}.

%% ################################################################# %%
\subsection{The energetics of Solar flares}\label{sol_flr_energ}
%% ################################################################# %%

Given the rich history of flare observations over more than 150 years, remarkably enough, we have barely started scratching the surface of the complexities involved in the physics of Solar flares. The reconfiguration of the magnetic structures, which almost always involves complex geometry, makes almost all events unique in some sense. Following the reconfiguration of the magnetic field, the released magnetic free energy is transported across various layers of the Sun and converted into various other forms of energy. The overarching conversion of magnetic energy into various forms, i.e., reconnection outflow, particle acceleration, thermal and kinetic energy of chromospheric plasma, etc, has proven incredibly complex to model. In cases where flares are associated with eruptions, e.g., CMEs, there is the added complexity of the kinetic and potential energy of the CMEs, the associated shocks and the kinetic energy of the Solar energetic particles.
%Considering how the energy is transformed the magnetic energy of the active region that is released after the reconnection into the reconnection outflow jets, the kinetic energy of escaping particles, the thermal and the kinetic energy of the Chromospheric plasma evaporating, the radiative and conductive losses. In case of the eruptive events, there is the added complexity of the kinetic and potential energy of the CMEs, the energy of the shocks and the kinetic energy of the solar energetic particles.

To constrain the models of Solar flares and the various associated phenomena, detailed quantitative characterization is absolutely necessary. %There have been several studies quantifying the partition between various subsets of the energies. 
The questions which are of particular importance are:

@@

%--------------------------------------------------------------------
\section{Motivation}\label{sec:mot}
%--------------------------------------------------------------------
%\info{This section is what we shall use as the synopsis. So, it should have : 1 st para: What are flares and why are they important. 2nd para: what is know and what is missing (only the most important aspects). 3rd Para: what is the overall goal, and what extra do you do? Just briefly. Thereafter Chapter wise summary.}

>>>>

The Sun is our nearest astrophysical object, which serves as a natural laboratory for various branches of Physics - atomic physics, spectroscopy, turbulence, magneto-hydrodynamics, stellar evolution, dynamo and magnetism and plays a key role in space and terrestrial weather. The Sun is also the primary source of light and energy on Earth. So it is only natural that we try to understand the different physical processes involved in the energetics of the Sun, which directly affects life on Earth. In this chapter, I briefly introduce the Sun, various layers of its atmosphere and solar flares. Finally, we provide a motivation for the thesis and a brief outline.

%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The Sun}\label{Sun}
%%%%%%%%%%%%%%%%%%%%%%%%%

The Sun is a G2V star, with surface luminosity $3.86~\times~10^{26}$~W and an effective temperature of $\sim 5780~K$. The Sun is mostly made out of Hydrogen (92.1\%) and Helium (7.8\%), along with negligible quantities of heavier elements C, N, O, Mg, Si, Ne and Fe, among others. The study of the Sun can be divided into three components, \textit{viz.} the solar interior, the solar atmosphere and the Heliosphere. Fig.~\ref{fig_solar_int} shows a schematic diagram of various regions of the Sun. Below, we briefly describe these three regions.

%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The Solar Interior}\label{solar_int} 
%%%%%%%%%%%%%%%%%%%%%%%%%

Depending on various physical characteristics, the solar interior can be divided into three zones {\it e.g.} the core, radiative zone and convective zone. In the innermost zone, the core, the Sun has a temperature of about 15~MK and a density of $\mathrm{1.6\times 10^{5}~km.m^{-3}}$. Here Hydrogen fuses into Helium primarily by p-p chain reaction \citep{bethe38} and to some extent by CNO cycle \citep{bethe39}. Beyond the core, we have the optically thick radiative zone, where the Hydrogen and Helium are completely ionized due to very high density. The high energy $\gamma$-ray photons generated in the core have a very small mean free path in this region, as they collide numerous times and propagate across this region. As we move radially outwards, temperature and density both decrease slowly due to gravitational equilibrium. Due to this reduction of temperature, there is a decrease in the thermal motion of the particles, and the hydrogen and helium start to recombine, giving rise to an increase in opacity at the top of the radiative zone. This change in opacity makes the propagation of photons via radiation less efficient, as photons get trapped and reabsorbed more. This creates a sharp temperature gradient due to the "photon pileup". This temperature gradient gives rise "convective instability". This is the convective zone where the energy from the inner layers of the Sun is transported adiabatically to the surface of the Sun.
%In between the radiative and the convective zone, there is \sroy{believed to be }a thin layer, roughly at $\sim~0.7~R_{\odot}$ known as the tachocline. \sroy{This region} is believed to play a major role in generating Sun's magnetic field via the Dynamo process \citep{corbard01}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The Solar Atmosphere}\label{solar_atmos}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The layers outside the convection zone together form the solar atmosphere and is further divided into different regions. The photosphere, which is also known as the Sun's surface, the chromosphere, the transition region and the corona. While we can define geometrical height or depths to define these layers, it is much more useful to define them using the local optical depths of various spectral lines, which are directly used to observe the layers. This also gives us an idea about which spectral features originate mainly from which portions of the solar atmosphere. Fig.~\ref{fig_solar_atm} shows the variation of temperature and number density across various layers of the solar atmosphere \citep{philips08}.

%% ############# %%
\begin{figure}[ht!]
    \centering
    \includegraphics[width = 0.8\linewidth]{Figures/solar_atm.png}
    \caption[Density and temperature profiles across various layers of the Sun.]{The variation of electron temperature (solid line) and electron density ($\mathrm{N_{e}}$, dashed line) and density of neutral hydrogen atoms ($\mathrm{N_{H}}$, dot-dashed line) in the solar atmosphere \citep{philips08}.}
    \label{fig_solar_atm}
\end{figure}
%% ############# %%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{The Photosphere}\label{photosphere}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The photosphere is defined to be the layer where $\tau_{\lambda~\sim~5500}=1$. This is in the green part of the visible spectrum and the Sun is opaque in visible beyond this layer, hence it is called to be the surface of the Sun. This layer is 400{--}600~km thick and has an effective temperature of $\sim$5780~K. The magnetic field lines arising from the tachocline penetrate the photosphere and create a `carpet' over the whole region \citep{priest14}. As mentioned earlier, this is the solar surface known as the `quiet Sun' (QS). This region exhibits an average magnetic flux density of 10{--}50 G. The QS surface is covered with cells of roughly four sizes, namely, granules, mesogranules, supergranules and giant cells.

Some regions exhibit much stronger magnetic flux density, often associated with highly twisted magnetic field structures. These magnetic features are manifested as Sunspots. The Sunspots harbour some of the strongest concentrated magnetic fields, reaching up to several kilogauss in strength. The other noteworthy solar magnetic structure apart from these are believed to be spatially unresolved flux tubes composing a network with field strengths $\sim$ {\it $O(10^{3} G)$}\citep{solanki93,grossman96,rubio19}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{The Chromosphere}\label{chromosphere}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The chromosphere is a highly non-uniform dynamic layer with a thickness of 1500{--}3000~km, with increasing temperature (up to $10^{4}$ K) and decreasing number density. As seen in Fig.~\ref{fig_solar_atm}, the temperature of the chromosphere saturates before the dramatic rise in the transition region. This saturation is attributed to a steady deposition of acoustic energy by the creation of shock waves \citep{carlsson07}. The chromosphere also exhibits a sharp gradient in the plasma $\beta$ factor, non-LTE conditions, and the dominance of wave motions. The observable features in the chromosphere are active regions, sunspots, plages and also light bridges as labelled in images taken by IRIS SJI (slit-jaw imager) 2796 {\AA} (left panel) and 2832 {\AA} (right panel) displayed in Fig.~\ref{fig:sji_features}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{The Corona}\label{corona}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The outermost layer of the atmosphere is the corona. The corona maintains a temperature of 1-2~MK during ambient conditions but can rise up to tens of MK during explosive eruptions such as solar flares. Due to such high temperatures, the coronal gas is fully ionized. It is only visible to the naked eye during the total solar eclipse or with coronagraphs due to Thomson scattering of the photospheric visible photons with free electrons in the corona. Corona also emits in various emission lines of e.g. Mg, Fe, Si, S, Ca, O, etc. and their ions, across a wide band of electromagnetic spectrum ranging from X-ray to visible wavelengths. Similar to the other layers, the corona exhibits a variety of structures such as ARs, coronal holes, bright points, filaments, and flares. The flare arcades are most prominently visible in the corona. %\info{I made a lot of changes in this section. please see if all is consistent.}

@@

solar flares are the most powerful energetic explosions in the solar system. They are described as a sudden increase in brightness in localized areas of the Sun. Within tens of minutes, they can release over $10^{32}$ erg of energy, which is emitted across the entire electromagnetic spectrum, i.e., from radio to gamma rays. They can also launch highly energetic particles into the interplanetary medium. Flares occur in magnetic active regions. It has been estimated that the amount of flare energy released is comparable to the free energy stored in the magnetic system \citep{emslie12,ash17}. The total energy released varies from event to event. It is also known that larger events occur much less frequently than smaller events.

Fig.~\ref{fig:sji_features} displays one of the most well-known and well-studied solar eruptions from Aug 31st, 2012. During this event, an associated giant prominence erupted from the southeast limb, which was visible in all AIA channels as shown in Fig.~\ref{fig:aia-feature}. The sunspots are also clearly visible and labelled in the HMI continuum (see Fig.~\ref{fig:aia-feature}h). In Fig.~\ref{fig:aia-feature}i, we show the co-aligned AIA 304 {\AA} and HMI continuum observation. The prominences are rooted at the sunspot penumbra, accompanied by the hot thermal plasma associated with the accompanying flare (visible in Fig.~\ref{fig:aia-feature} e {--} f). The novelty of such observation lies in the clear spatial connection across various layers of the Sun, connected via energy and momentum transported from one layer to the other.

@@

On September 1, 1859, R.C. Carrington and R. Hodgson independently observed the first flare in white-light continuum \citep{carrington1859, hodgson1859}. Such localized brightening on the Sun has remained an enigma ever since. Shortly after the observation by Carrington and Hodgson, the Sun was being studied extensively in the H$\alpha$ wavelength, which essentially images the chromosphere, and the reports of flaring events became increasingly frequent and progressively more complex. No two events were similar, as there were variations observed in the size of the source, ejections of plasma, and shock waves driving into the interplanetary space. Advances in radio technology during the Second World War affirmed the detection of the presence of non-thermal electrons in the solar corona during military radar operations \citep{hey46}. Around the same time, S.E. Forbrush noticed ground-level cosmic ray enhancement during major solar flares. These discoveries eluded that the flaring events do not only involve the thermal plasma but somehow connect with high energy particles and involve the corona.

In the late 1950s, using rockets and balloons, the Sun was observed in hard X-rays ($\ge10~keV$). \cite{peterson59} discovered the first hard X-ray emission during a flare in 1958. Later on, it was deduced from the enhancements observed in radio and hard X-rays that the ejected energetic particles may contain a substantial fraction of the initial energy released \citep{brown71}. Note that while the hard X-ray is created by the bremsstrahlung radiation of the electrons colliding into dense material, resulting in a power-law energy distribution, the broadband radio emission from 1 to 100 GHz is created from gyrosynchrotron emission. Flares are now routinely observed with different space-borne instruments, in particular in the EUV and soft X-ray($\le$~10~keV), which shows that the energy released from the flare heats the local plasma to temperatures beyond 30~MK.

@@

\noindent and is known as the ``Neupert effect''. This relationship shows that the soft X-ray radiation primarily originates from a thermal plasma which is heated by the energy of the flaring event deposited by the accelerated electrons. We note, however, that Eqn.~\ref{npt_eff_eq} is only valid if cooling by conduction is negligible and/or the role of ions in the flare energy deposition is minimal \citep{veronig05}.

%% ################################################################# %%
\subsection{``Standard'' model of solar flare}\label{sol_flr_std_mod}
%% ################################################################# %%

Based on various multi-wavelength studies of solar flares, combined with theoretical modelling a `standard' model of solar flares has been advocated by \cite{carmichael64,sturrock66,hirayama74,kopp76}, and termed as \textsl{CSHKP} model. The flares occur due to the abrupt release of magnetic free energy, which was previously stored in the coronal magnetic field via flux emergence and surface motions \citep{forbes06}. The majority of the strongest flares are eruptive in nature. The `standard' model attributes the release of flare energy to magnetic reconnection in the corona, which happens due to the eruption and/or coronal mass ejection \citep{shibata95,lin_n_forbes00,moore01,priest_n_forbes02}. It is noteworthy that, despite continuous attempts over the decades, this is not at all a quantitative model. Rather, it is an attempt to unify the multi-wavelength observations and modelling attempts in an orderly manner. Naturally, the model can not explain all the observations from various flares. A schematic picture of the model is shown in Fig.~\ref{fig:std_mod}. Below we describe the salient points of the `standard' model:

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[ht!]
    \centering
    \hspace{5cm}
    \includegraphics[trim={1cm 8cm 0cm 8cm}, clip, width=0.95\linewidth]{Figures/std_mod.pdf}
    \caption{A schematic diagram of the standard model of solar flare.}
    \label{fig:std_mod}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%
\begin{enumerate}
    \item In the standard model, following the reconnection event that releases the magnetic free energy, electrons and ions from the reconnection site are accelerated along the realigned magnetic field lines downwards. In the case of an associated CME, parts of the magnetic energy is deposited as kinetic energy. The coronal reconnection source where the energy is released is often observed in HXR. The loop-top HXR sources are often accompanied by the thermal SXR sources. The loop top HXR source is observed above the coronal SXR source \citep{masuda94}.

    \item The accelerated particles move along the field lines downwards along the magnetic field lines, giving rise to the microwave and radio observations seen from flares via gyro magnetic radiation at various frequencies dictated by the characteristics of local plasma. Finally, the accelerated particles hit the chromosphere, which is considerably denser than the corona, giving rise to Hard X-rays and sometimes also gamma-rays observed from the foot points. This also explains why the coronal hard X-ray source is considerably softer than the foot points. The energetic particles deposit their energy into the local chromosphere as they go through a series of collisions and eventually thermalize \citep{brown83}. Foot point sources are occasionally detected on H-alpha flare ribbons marking the base of a magnetic arcade \citep{masuda01}. The ribbon brightening is usually observed to propagate along the ribbons while they separate further from each other, signifying continued reconnection \citep{tripathi06}.

    \item $\gamma$-ray emission from solar flares is visible between 0.8 {--} 20 MeV generated by the accelerated ions. The neutron capture line at 2.2 MeV forming Deuterium is usually the strongest characteristic line visible from such emissions. Flare-accelerated protons and ions colliding with the dense chromosphere can produce neutrons. It is suggested that these neutrons are captured by ambient protons to form Deuteron, giving rise to the 2.2 MeV characteristic bump \citep{ramtay74,hua87}. Later, RHESSI observations \citep{hurford_2003,hurford_2006} showed that the $\gamma$-ray foot-points did not coincide with the HXR foot-points. This demonstrated that the ions and electrons were accelerated differently, as proposed by \cite{emslie_2004}.

    \item  The thermalized plasma emits in thermal bremsstrahlung, giving rise to soft X-ray and EUV radiation. The heating of the local chromosphere increases the local pressure. When the growing pressure gradient builds up enough, the local plasma expands upwards and fills up the coronal loops with soft X-ray-emitting plasma. This phenomenon is known as "chromospheric evaporation". This was first reported by \cite{doscheck80,feldman80}. \cite{antonucci89} reported velocities of 300 {--} 400 km/s with temperature $\sim$ 20 MK in \ion{Ca}{19}, while later on \cite{milligan06} reported much gentler up-flow with velocities $\sim$ 200 km/s in \ion{Fe}{19} line from {\it SOHO}/CDS observations. Several simulations also reproduced evaporation arising from non-thermal particle precipitation \citep{sterling93,hori98,reeves07}.
\end{enumerate}
%%%%%%%%%%%%%%%%%

It is important to emphasise that while the standard model attempts to explain and unify various kinds of differences seen from the numerous flares we observe, there are several cases where the standard model cannot explain the observations. For example, there have been observations of flares where the hard X-ray footpoint does not form \citep{veronig02,veronig04}. It is proposed that in these cases, the plasma in the coronal loops is so dense that the accelerated particles pass through enough collisions to thermalize within the flare loops before reaching the chromosphere. In this process, they may defuse the energy more evenly within the loop rather than impulsively dumping it at the base of the loops. Another flare which previously occurred at the same region might explain the denser coronal loops \citep{strong84,bone07}.

%% ################################################################# %%
\section{Motivation of the thesis: the energetics of solar flares}\label{sol_flr_energ}
%% ################################################################# %%

Given the rich history of flare observations over more than 150 years, we have learnt a lot about their charactershtics and dyamics. However, there are several complexities involved in the physics of solar flares, which are still fully understood. The reconfiguration of the magnetic structures, which almost always involves complex geometry, makes almost all events unique in some sense. Following the reconfiguration of the magnetic field, the released magnetic free energy is transported across various layers of the Sun and converted into various other forms of energy. The overarching conversion of magnetic energy into various forms, i.e., reconnection outflow, particle acceleration, thermal and kinetic energy of chromospheric plasma, etc., has proven incredibly complex to model. In cases where flares are associated with eruptions, e.g., CMEs, there is the added complexity of the kinetic and potential energy of the CMEs, the associated shocks and the kinetic energy of the solar energetic particles.
%Considering how the energy is transformed the magnetic energy of the active region that is released after the reconnection into the reconnection outflow jets, the kinetic energy of escaping particles, the thermal and the kinetic energy of the Chromospheric plasma evaporating, the radiative and conductive losses. In case of the eruptive events, there is the added complexity of the kinetic and potential energy of the CMEs, the energy of the shocks and the kinetic energy of the solar energetic particles.

To constrain the models of solar flares and the various associated phenomena, detailed quantitative characterization is absolutely necessary. %There have been several studies quantifying the partition between various subsets of the energies. 
The questions which are of particular importance are:

@@

%--------------------------------------------------------------------
%\section{Motivation}\label{sec:mot}. 
%--------------------------------------------------------------------
%\info{This section is what we shall use as the synopsis. So, it should have : 1 st para: What are flares and why are they important. 2nd para: what is know and what is missing (only the most important aspects). 3rd Para: what is the overall goal, and what extra do you do? Just briefly. Thereafter Chapter wise summary.}
